---
title: "DATA607-Project 1"
author: "Juanelle Marks"
date: "9/23/2018"
output: html_document
---

##Project Question

>In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605
1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.


```{r}
library(stringr)
library(knitr)
library(dplyr)

```


###Sourcing Data
```{r}
tournamentinfo<-"https://raw.githubusercontent.com/juanellemarks/JuanelleMarks_DATA607_2018/master/tournamentinfo.text"

textFile <- read.csv(tournamentinfo, header = FALSE)

textFile[c(1:10), ]
```


###Data Cleansing
```{r}
#remove first four rows as they are unecessary to the solution
textFile1<-textFile[-c(1:4),]
head(textFile1,3)
```


```{r}
#Convert from factor to character
textFile1 <- as.character(textFile1)
head(textFile1,3)
```


##Data Manipulation
```{r}
#extract rows with player information
playerInfo<-textFile1[seq(1, length(textFile1), 3)] 
head(playerInfo,3)
```


```{r}
#extract rows with rating information
ratingInfo<-textFile1[seq(2, length(textFile1), 3)]
head(ratingInfo,3)
```


```{r}
#remove unwanted characters from rows
player_id <- as.integer(str_extract(playerInfo, "\\d+"))

playerName <- str_trim(str_extract(playerInfo, "(\\w+\\s){2,3}"))

player_state <- str_extract(ratingInfo, "\\w\\w")
totalPoints<- as.numeric(str_extract(playerInfo, "\\d.\\d"))

pre_Rating<- as.integer(str_extract(str_extract(ratingInfo, "\\D\\d{3,4}\\D"), "\\d+")) 

opponent <- str_extract_all(str_extract_all(playerInfo, "\\d+\\|"), "\\d+")

```



```{r}
opponent_count<- sapply(opponent, length)
seq.max <- seq_len(max(opponent_count))
z <- t(sapply(opponent, "[", i = seq.max)) 
textFile2 <- as.data.frame(z)


y <- data.frame(player_id, pre_Rating)
textFile3 <- textFile2
textFile3[] <- y$pre_Rating[match(unlist(textFile3), y$player_id)]
kable(head(textFile3))

#  Finding average pre-Chess rating of opponents
textFile3$oppAvg <- round(apply(textFile3, 1, mean, na.rm=TRUE))
kable(head(textFile3))
```



```{r}
final_textFile <- data_frame(player_id, playerName, player_state , totalPoints, pre_Rating, textFile3$oppAvg )

colnames(final_textFile) <-  c("ID", "Player_Name", "Player_State", "Total_Points", "Pre_Match_Rating", "Opponent's_Average")

kable(head(final_textFile, 10))
```

## Write to CSV File
```{r}
write.csv(final_textFile,file="/Users/juanelle/Desktop/MSDS/Data607/Week Three/Project 1/DATA607_ProjectOne_Newtxtfile")
```







