---
title: "DATA607-Assignment3"
author: "Juanelle Marks"
date: "9/16/2018"
output: html_document
---

```{r}
library(stringr)
```

##Question 3

####Part A
> Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

```{r}
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555
-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson,Homer5553642Dr. Julius Hibbert"
```


```{r}
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```



```{r}
# Change order
name2 <- str_replace_all(name, "(.+)(, .+)$", "\\2 \\1") 
name2 
```

```{r}
# Remove commas
name3 <- str_replace_all(name2, "," , " ") 
name3
```


```{r}
#Remove titles
no_title <- str_replace(name3, "[[:alpha:]]{2,}\\. ", "")
no_title 
```


####Part B

>Construct a logical vector indicating whether a character has a title (i.e., Rev. andDr.).

```{r}
title_Check <- c(str_detect(name, "(^[A-z]{1,}\\. ).*"))
title_Check <- data.frame(name, title_Check)
names(title_Check) <- c("Name", "Has Title?")
title_Check
```

#####Part C

>Construct a logical vector indicating whether a character has a second name.

```{r}
secondname_Check <- str_detect(name, " [A-z]{1,}\\. ")
secondname_Check <- data.frame(name, secondname_Check)
names(secondname_Check) <- c("Name", "Has Second Name?")
secondname_Check
```

##Question 4

#####Part A

>[0-9]+\\$ 
>Strings that confirm to the above pattern are those that contain a series of numbers from 0 to 9 and ends with the $ sign.

#####Example:
```{r}
example1 <- c("Juanelle1978$Anthony", "$123456", "987654321$123", "6789$", "$23sf")
example1 <- unlist(str_extract_all(example1, pattern = "[0-9]+\\$" ))
example1
```


#####Part B

 >\\b[a-z]{1,4}\\b
 > Strings that confirm to the above pattern are those that  matches one to four-character words in the input string ( the four character words words must not contain punctuation marks or digits and must be  lowercase).

#####Example:
```{r}
example2 <- c("Timothy", "i am a  boy", "123 abc doh rey me", "891 abcd efghi", "ABC abc XYZ xyz")
example2 <- str_extract_all(example2, pattern = "\\b[a-z]{1,4}\\b" )
example2
```

####Part C
>.*?\\.txt$
> Strings that confirm to the above pattern are those that end with .txt. The above pattern is great for checking file types.

#####EXample:

```{r}
example3<- c("Juanelle.txt", "911.txt",".txtMarks", "1234.txt", "Anthony Marks xyz.txt")
example3 <- str_extract_all(example3, pattern = ".*?\\.txt$" )
example3
```


#####Part D
> \\d{2}/\\d{2}/\\d{4}
>Strings that confirm to the above pattern are those that begin with two digits  immediately followed by a forward slash then immediately followed by two digits and a forward slash, then immediately followed by four digits. The output resembles a date format. Great for checking for the presence of dates.

#####Example
```{r}
example4 <- c("09/12/2018", "ab/cd/ efgh", "09/ab/2018", "111/2/19654")
example4 <- str_extract_all(example4, pattern = "\\d{2}/\\d{2}/\\d{4}")
example4
```


#####Part E
> <(.+?)>.+?</\\1>
>Strings that confirm to the above pattern are those that confirm to the same  pattern as open and close html tags.


#####Example:
```{r}
example5<-c("<p> This is a paragraph</p>", "<p> This is a paragraph</z>", "<p> This is a paragraph", "This is a paragraph</p>", "<h1> This is a paragraph</h1>", "123abc")
example5 <- str_extract_all(example5, pattern = "<(.+?)>.+?</\\1>")
example5
```


##Question 9
> The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others!

>"clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"


>A message was detected when all uppercase characters were extracted from the code. See below:

```{r}
code<-"clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

secret_message<-unlist(str_extract_all(code, "[[:upper:].]{1,}"))

secret_message <-  str_replace_all(string =  secret_message, pattern =  "\\.", replacement = " ")

paste(secret_message, collapse = "")

```

